---
title: "Chapter 5 Exercises"
author: "J.Chapman"
date: "9/20/2019"
output: html_document
---
```{r setup, include=FALSE}
library(nycflights13)
library(tidyverse)

```
# Exercises 5.2.1

## 1. all flights with an arrival delay of 120 or more

```{r echo=FALSE}
filter(flights, arr_delay >= 120)
```
total = 10,200 flights (rows)

## 2. all flights that flew to Houston (IAH or HOU)
```{r echo=FALSE}
filter(flights, dest == "IAH" | dest == "HOU")

```
total rows = 9,313

## 3. all flights operated by United, American, or Delta airlines
DL = Delta, AA = American, UA = United
```{r echo=FALSE}
filter(flights, carrier %in% c("UA", "AA", "DL"))

```
total rows = 139,504 (total is 336,776), so 139504/336776 = 41.42%

## 4. all flights that departed in the summer (july, aug, sep)
```{r echo=FALSE}
filter(flights, month %in% c("7","8","9"))
# or
# filter(flights, month >=7, month <=9)
# total rows = 86,326
# or
# filter(flights, month %in% 7:9)
# total rows = 86,326
# or
# (less efficient, more code)
# filter(flights, month == 7 | month == 8 | month == 9)
# total rows = 86,326

```
total rows = 86,326

## 5. all flights that arrived more than two hours late, but didn't leave late
```{r echo=FALSE}
filter(flights, dep_delay <= 0 & arr_delay > 120)
# or
# filter(flights, arr_delay > 120, dep_delay <= 0)
# total rows = 29, or 29/336776 = .009% of flights

```
total rows = 29


## 6. all flights delayed by at least an hour, but made up over 30 minutes in flight
```{r echo=FALSE}
filter(flights, dep_delay >= 60 & dep_delay - arr_delay > 30)

```
total rows = 1844

## 7. Flights that departed between midnight and 6 a.m.
```{r echo=FALSE}
# Finding flights that departed between midnight and 6 a.m. is complicated by the way in which times are represented in the data.
# In dep_time, midnight is represented by 2400, not 0. You can verify this by checking the minimum and maximum of dep_time.
# 
# summary(flights$dep_time)
# #>    Min. 1st Qu.  Median    Mean 3rd Qu.    Max.    NA's 
# #>       1     907    1401    1349    1744    2400    8255
# This is an example of why it is always good to check the summary statistics of your data. Unfortunately, this means we cannot simply check that dep_time < 600, because we also have to consider the special case of midnight.

filter(flights, dep_time <= 600 | dep_time == 2400)

```
total rows = 9373

# 5.2.3 How many flights have a missing dep_time? What other variables are missing? What might these rows represent?
```{r echo=FALSE}
# Find the rows of flights with a missing departure time (dep_time) using the is.na() function.
filter(flights, is.na(dep_time))
```



